They're Clutching up! Team Momentum in Round-Based Esports

### 2022-08-14 paper2/paper2.Rmd
try 2

---
title: "They're Clutching up! Team Momentum in Round-Based Esports"
authors:
  - name: Tony ElHabr
    affiliations:
      - name: Georgia Institute of Technology
    email: anthonyelhabr@gmail.com
    url: https://tonyelhabr.rbind.io
abstract: |
  My research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD).

  First, I find evidence in CoD defying the naive hypothesis that a series represents a sequence of independent events (rounds), with each team having a constant 50% probability of winning a given round.

  Second, I examine post-streak round win probability. I find that teams perform significantly worse than expected after streaks of 2, 3, and 4 wins when series end up going to 9, 10, or 11 (maximum) rounds, even after accounting for the "hot-hand" phenomenon.
bibliography: references.bib  
biblio-style: unsrt
output: rticles::arxiv_article
---

@@

```{r}
#| label: setup-code
library(dplyr)
library(qs)
library(tidyr)
library(gt)
library(scales)

@@

The empirical offensive round win percentage across all rounds is 47.8%.[^6] To contextualize my subsequent discussion of series length and momentum, table \ref{tbl:cod-o-win-prop-by-series-state} shows round win percentages by series "state" (i.e. the number of round wins by each team prior to an upcoming round). Offensive round win rate is not quite constant, although never veers more than 10% from this global average.

@@

Despite the plethora of existing research on games played in a series and momentum in sports, these topics have yet to be investigated heavily in esports. Work has been done to examine in-round win probability in other FPS titles such as Counter-Strike ([@xenopoulos2022]) and Valorant ([@derover2021]), both of which are round-based like CoD SnD. However, research on round-level trends is sparse, perhaps because games like Counter-Strike and Valorant both have economic aspects that can create clear advantages on side in a given round, given how prior rounds played out.[^9]

@@

Table \ref{tbl:cod-prob-series-lasting-i-rounds} shows the expected proportion for $s = 11$ and the naive assumption that $p = 0.5$, along with the observed frequencies observed in CoD SnD, as both a count $N_O(i)$ (of series lasting $i$ rounds) and a proportion $P_O(i)$ of all series.

```{=tex}
\begin{longtable}{rrrr}
  \caption{The probabilities that a best-of-11 series lasts $i$ rounds ($Pr(i)$, where $i = {6, 7, 8, 9, 10, 11}$) under the assumption that each team has a 50% probability ($p = 0.5$) of winning each round.}\label{tbl:cod-prob-series-lasting-i-rounds} \\
  \toprule
  &  & \multicolumn{2}{c}{Observed} \\ 
  \cmidrule(lr){3-4}
  Series lasts $i$ rounds & $P_E(i)$ & $P_O(i)$ & $N_O(i)$ \\ 
  \midrule
  6 & $3.1\%$ & $4.7\%$ & 40 \\ 
  7 & $9.4\%$ & $11.9\%$ & 101 \\ 
  8 & $16.4\%$ & $16.5\%$ & 141 \\ 
  9 & $21.9\%$ & $21.7\%$ & 185 \\ 
  10 & $24.6\%$ & $21.5\%$ & 183 \\ 
  11 & $24.6\%$ & $23.7\%$ & 202 \\ 
  \bottomrule
\end{longtable}
```
The probabilities that a best-of-11 series lasts $i$ rounds ($P(i)$, where $i \in R = [6, 7, 8, 9, 10, 11]$) under the assumption that each team has a 50% probability ($p = 0.5$) of winning each game {#tbl-prob-series-lasting-i-rounds} along with the observed frequencies of series lasting $i$ rounds, expressed as a percentage $P_O(i)$ and as a count ($N_O$)

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq-series-length}{}{
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(i) - P_E(i))^2}{P_E(i)}, i \in R = [6, 7, 8, 9, 10, 11].
}\label{eq:chi-squ-test}\end{equation}
```
as 16.0 (p-value of 0.0068), I can comfortably reject the constant probability null hypothesis, event at a confidence level of $\alpha = 0.01$.

@@

cod_streaks_other_res |>
  mutate(
    lab = sprintf('%s. p = %s', idx, p)
  ) |>
  select(lab, series_streaks) |>
  unnest(series_streaks) |>
  select(lab, n_rounds, prop_expected) |>
  mutate(
    across(prop_expected, ~scales::percent(.x, accuracy = 0.1))
  ) |>
  pivot_wider(
    names_from = lab,
    values_from = prop_expected
  ) |>
  inner_join(
    cod_streaks_naive_res$series |>
      transmute(n_rounds, `$P_O(i)$` = scales::percent(prop_actual, accuracy = 0.1)),
    by = 'n_rounds'
  ) |>
  arrange(n_rounds) |>
  rename(`Series lasts $i$ rounds` = n_rounds) |>
  knitr::kable()
```

@@

```{=tex}
\begin{longtable}[]{@{}lrr@{}}
  \caption{Alternate estimates of the constant probability \(p\) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive \(p = 0.5\).}\label{tbl:mosteller-methods-results} \\
  \toprule()
    Method & \(p\) & \(\chi^2\) (p-value) \\
    \midrule()
    \endfirsthead
    \toprule()
    Method & \(p\) & \(\chi^2\) (p-value) \\
    \midrule()
  \endhead
  0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
  1. Method of moments & 0.5725 & 3.6 (0.6) \\
  2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
  3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
  \bottomrule()
\end{longtable}
```
Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new $P_E(i)$ when re-applying Equation \ref{eq:series-length} for each new $p$. I observe that $P_E(i)$ is notably smaller for $p_{1,2,3}$ when $i \in [10, 11]$ and higher for $i \in [6, 7]$, more closely matching $P_O(i)$.

```{=tex}
\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2427}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1456}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1748}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1650}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1748}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0971}}@{}}
\caption{The probabilities that a best-of-11 series lasts $i$ rounds under various assumptions for $p$.}\label{tbl:expected-series-lengths-alternative-ps} \\
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Series lasts \(i\) rounds
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
0. \(p\) = 0.5
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
1. \(p\) = 0.5725
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
2. \(p\) = 0.575
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
3. \(p\) = 0.5775
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(P_O(i)\)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Series lasts \(i\) rounds
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
0. \(p\) = 0.5
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
1. \(p\) = 0.5725
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
2. \(p\) = 0.575
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
3. \(p\) = 0.5775
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(P_O(i)\)
\end{minipage} \\
\midrule()
\endhead
6 & 3.1\% & 4.1\% & 4.2\% & 4.3\% & 4.7\% \\
7 & 9.4\% & 11.1\% & 11.2\% & 11.4\% & 11.9\% \\
8 & 16.4\% & 17.7\% & 17.8\% & 17.9\% & 16.5\% \\
9 & 21.9\% & 21.8\% & 21.8\% & 21.8\% & 21.7\% \\
10 & 24.6\% & 23.1\% & 23.0\% & 22.9\% & 21.5\% \\
11 & 24.6\% & 22.1\% & 22.0\% & 21.8\% & 23.7\% \\
\bottomrule()
\end{longtable}
```
So I conclude that the constant round win probability assumption can be valid in CoD SnD series with the appropriate choice of $p$ ($\approx 0.575$).

>>>>

---
title: "They're Clutching up! Team Momentum in Round-Based Esports"
authors:
  - name: Tony ElHabr
    affiliations:
      - name: Georgia Institute of Technology
    email: anthonyelhabr@gmail.com
    url: https://tonyelhabr.rbind.io
abstract: |
  My research investigates patterns in round win percentages in professional Search and Destroy (SnD) matches of the popular first-person shooter game Call of Duty (CoD). I find evidence in CoD supprting the hypothesis that round win probability can be modeled as constant across the series, although not at a naive 50%. Second, I examine post-streak round win probability, in search of evidence positive recency (the "hot hand" fallacy) or negative recency (the "gambler's falacy"). I find that teams perform significantly worse than expected after streaks of 2, 3, and 4 wins when series end up going to 9, 10, or 11 (maximum) rounds, suggesting the presence of negative recency.
bibliography: references.bib  
output:
  rticles::arxiv_article:
    includes:
      in_header: preamble.tex
---

@@

```{r}
#| label: setup-code
library(dplyr)
library(qs)
library(tidyr)
library(purrr)
library(gt)
library(scales)

@@

The empirical offensive round win percentage across all rounds is 47.8%.[^6] Table \ref{tbl:cod-o-win-prop-by-series-state} shows round win percentages by series "state" (i.e. the number of round wins by each team prior to an upcoming round). Offensive round win rate is not quite constant, although never veers more than 10% from this global average.

@@

Despite the plethora of existing research on games played in a series and momentum in sports, these topics have yet to be investigated heavily in esports. Work has been done to examine in-round win probability in other FPS titles such as Counter-Strike [@xenopoulos2022] and Valorant [@derover2021], both of which are round-based like CoD SnD. However, research on round-level trends is sparse, perhaps because games like Counter-Strike and Valorant both have economic aspects that can create clear advantages on side in a given round, given how prior rounds played out.[^9]

@@

Table \ref{tbl:cod-prob-series-lasting-i-rounds} shows the expected probabilities for $s = 11$ under the naive assumption that $p_0 = 0.5$ for a given round, along with the observed frequencies observed in CoD SnD ($P_O(i)$, $N_O(i)$)).

```{=tex}
\begin{longtable}{rrrr}
  \caption{The probabilities that a best-of-11 series lasts $i$ rounds ($P_E(i)$, where $i = {6, 7, 8, 9, 10, 11}$) under the assumption that each team has a 50% probability ($p_0 = 0.5$) of winning each game. The observed frequencies observed are shown as a count $N_O(i)$ and as a proportion $P_O(i)$ of all series.}\label{tbl:cod-prob-series-lasting-i-rounds} \\
  \toprule
  &  & \multicolumn{2}{c}{Observed} \\ 
  \cmidrule(lr){3-4}
  Series lasts $i$ rounds & $P_E(i)$ & $P_O(i)$ & $N_O(i)$ \\ 
  \midrule
  6 & $3.1\%$ & $4.7\%$ & 40 \\ 
  7 & $9.4\%$ & $11.9\%$ & 101 \\ 
  8 & $16.4\%$ & $16.5\%$ & 141 \\ 
  9 & $21.9\%$ & $21.7\%$ & 185 \\ 
  10 & $24.6\%$ & $21.5\%$ & 183 \\ 
  11 & $24.6\%$ & $23.7\%$ & 202 \\ 
  \bottomrule
\end{longtable}
```
The probabilities that a best-of-11 series lasts $i$ rounds ($P(i)$, where $i \in R = [6, 7, 8, 9, 10, 11]$) under the assumption that each team has a 50% probability ($p_0 = 0.5$) of winning each game {#tbl-prob-series-lasting-i-rounds} along with the observed frequencies of series lasting $i$ rounds, expressed as a percentage $P_O(i)$ and as a count ($N_O$)

@@

```{=tex}
\begin{equation}\protect\hypertarget{eq-series-length}{}{
\chi^2 = \sum^{11}_{i=6} \frac{(P_O(i) - P_E(i))^2}{P_E(i)}, i \in R = [6, 7, 8, 9, 10, 11].
}\label{eq:chi-squ-test}\end{equation}
```
as 16.0 (p-value of 0.0068), I can comfortably reject the constant probability null hypothesis for $p_0 = 0.5$, event at a confidence level of $\alpha = 0.01$.

@@

cod_streaks_other_res |>
  filter(idx %in% c(0L, 2L)) |> 
  mutate(
    lab = sprintf('$p$ = %s', p)
  ) |>
  select(lab, series_streaks) |>
  unnest(series_streaks) |>
  select(lab, n_rounds, prop_expected) |>
  mutate(
    across(prop_expected, ~scales::percent(.x, accuracy = 0.1))
  ) |>
  pivot_wider(
    names_from = lab,
    values_from = prop_expected
  ) |>
  inner_join(
    cod_streaks_naive_res$series |>
      transmute(n_rounds, `$P_O(i)$` = scales::percent(prop_actual, accuracy = 0.1)),
    by = 'n_rounds'
  ) |>
  arrange(n_rounds) |>
  rename(`Series lasts $i$ rounds` = n_rounds) |>
  knitr::kable()
```

@@

```{=tex}
\begin{longtable}[]{@{}lrr@{}}
  \caption{Alternate estimates of the constant probability \(p\) for winning a given round in a CoD SnD, applying the three methods suggested by Mosteller (1952), in addition to the naive \(p = 0.5\).}\label{tbl:mosteller-methods-results} \\
  \toprule()
    Method & \(p\) & \(\chi^2\) (p-value) \\
    \midrule()
    \endfirsthead
    \toprule()
    Method & \(p\) & \(\chi^2\) (p-value) \\
    \midrule()
  \endhead
  0. Naive & 0.5000 & 16.0 (\textless=0.01) \\
  1. Method of moments & 0.5725 & 3.6 (0.6) \\
  2. Maximum likelihood & 0.5750 & 3.5 (0.62) \\
  3. Minimum \(\chi^2\) & 0.5775 & 3.5 (0.62) \\
  \bottomrule()
\end{longtable}
```
Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new $P_E(i)$ when re-applying Equation \ref{eq:series-length} for each the maximum likelihood estimate $p_2 = 0.575$.[^11] I observe that $P_E(i)$ is notably larger for $p_{2}$ when $i \in [6, 7]$, more closely matching $P_O(i)$. $p_{2}$ is also closer to $P_O(i)$ for $i \in [9, 10]$, although not for $i \in [8, 11]$.

[^11]: The method of moments and minimum $\chi^2$ estimates for $p$ are omitted simply because the results would be nearly identical to those for the maximum likelihood estimate of $p$ (since they are all $\approx 0.575$.

```{=tex}
\begin{longtable}[]{@{}rrrr@{}}
\caption{The probabilities that a best-of-11 series lasts $i$ rounds under the naive assumption $p_0 = 0.5$ and the maximum likelihood estimation $p_2 = 0.575$.}\label{tbl:expected-series-lengths-alternative-ps} \\
\toprule()
Series lasts \(i\) rounds & \(p_0\) = 0.5 & \(p_2\) = 0.575 & \(P_O(i)\) \\
\midrule()
\endhead
6 & 3.1\% & 4.2\% & 4.7\% \\
7 & 9.4\% & 11.2\% & 11.9\% \\
8 & 16.4\% & 17.8\% & 16.5\% \\
9 & 21.9\% & 21.8\% & 21.7\% \\
10 & 24.6\% & 23.0\% & 21.5\% \\
11 & 24.6\% & 22.0\% & 23.7\% \\
\bottomrule()
\end{longtable}
```
So I conclude that the constant round win probability assumption can be valid in CoD SnD series with the appropriate choice of $p$ ($\approx 0.575$).

### 2022-08-14 paper2/paper2.tex
try 2

\begin{document}
\maketitle

\begin{abstract}
My research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD).

First, I find evidence in CoD defying the naive hypothesis that a series
represents a sequence of independent events (rounds), with each team
having a constant 50\% probability of winning a given round.

Second, I examine post-streak round win probability. I find that teams
perform significantly worse than expected after streaks of 2, 3, and 4
wins when series end up going to 9, 10, or 11 (maximum) rounds, even
after accounting for the ``hot-hand'' phenomenon.
\end{abstract}

@@

The empirical offensive round win percentage across all rounds is
47.8\%.\footnote{Offensive round win percentage has been nearly constant
  across the three games during the CDL era: 1. 47.2\% in MW (2020) 2.
  47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)} To
contextualize my subsequent discussion of series length and momentum,
table \ref{tbl:cod-o-win-prop-by-series-state} shows round win
percentages by series ``state'' (i.e.~the number of round wins by each
team prior to an upcoming round). Offensive round win rate is not quite
constant, although never veers more than 10\% from this global average.

@@

Despite the plethora of existing research on games played in a series
and momentum in sports, these topics have yet to be investigated heavily
in esports. Work has been done to examine in-round win probability in
other FPS titles such as Counter-Strike ((Xenopoulos, Freeman, and Silva
2022)) and Valorant ((DeRover 2021)), both of which are round-based like
CoD SnD. However, research on round-level trends is sparse, perhaps
because games like Counter-Strike and Valorant both have economic
aspects that can create clear advantages on side in a given round, given
how prior rounds played out.\footnote{Additionally, both Counter-Strike
  and Valorant have overtime rules and blocked offensive/defensive roles
  (i.e.~playing either offense or defense for many consecutive rounds).}

@@

Table \ref{tbl:cod-prob-series-lasting-i-rounds} shows the expected
proportion for \(s = 11\) and the naive assumption that \(p = 0.5\),
along with the observed frequencies observed in CoD SnD, as both a count
\(N_O(i)\) (of series lasting \(i\) rounds) and a proportion \(P_O(i)\)
of all series.

\begin{longtable}{rrrr}
  \caption{The probabilities that a best-of-11 series lasts $i$ rounds ($Pr(i)$, where $i = {6, 7, 8, 9, 10, 11}$) under the assumption that each team has a 50% probability ($p = 0.5$) of winning each round.}\label{tbl:cod-prob-series-lasting-i-rounds} \\
  \toprule
  &  & \multicolumn{2}{c}{Observed} \\ 
  \cmidrule(lr){3-4}
  Series lasts $i$ rounds & $P_E(i)$ & $P_O(i)$ & $N_O(i)$ \\ 
  \midrule
  6 & $3.1\%$ & $4.7\%$ & 40 \\ 
  7 & $9.4\%$ & $11.9\%$ & 101 \\ 
  8 & $16.4\%$ & $16.5\%$ & 141 \\ 
  9 & $21.9\%$ & $21.7\%$ & 185 \\ 
  10 & $24.6\%$ & $21.5\%$ & 183 \\ 
  11 & $24.6\%$ & $23.7\%$ & 202 \\ 
  \bottomrule
\end{longtable}

The probabilities that a best-of-11 series lasts \(i\) rounds (\(P(i)\),
where \(i \in R = [6, 7, 8, 9, 10, 11]\)) under the assumption that each
team has a 50\% probability (\(p = 0.5\)) of winning each game
\{\#tbl-prob-series-lasting-i-rounds\} along with the observed
frequencies of series lasting \(i\) rounds, expressed as a percentage
\(P_O(i)\) and as a count (\(N_O\))

@@

as 16.0 (p-value of 0.0068), I can comfortably reject the constant
probability null hypothesis, event at a confidence level of
\(\alpha = 0.01\).

@@

Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new
\(P_E(i)\) when re-applying Equation \ref{eq:series-length} for each new
\(p\). I observe that \(P_E(i)\) is notably smaller for \(p_{1,2,3}\)
when \(i \in [10, 11]\) and higher for \(i \in [6, 7]\), more closely
matching \(P_O(i)\).

\begin{longtable}[]{@{}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2427}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1456}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1748}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1650}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1748}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0971}}@{}}
\caption{The probabilities that a best-of-11 series lasts $i$ rounds under various assumptions for $p$.}\label{tbl:expected-series-lengths-alternative-ps} \\
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Series lasts \(i\) rounds
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
0. \(p\) = 0.5
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
1. \(p\) = 0.5725
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
2. \(p\) = 0.575
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
3. \(p\) = 0.5775
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(P_O(i)\)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\raggedleft
Series lasts \(i\) rounds
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
0. \(p\) = 0.5
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
1. \(p\) = 0.5725
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
2. \(p\) = 0.575
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
3. \(p\) = 0.5775
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
\(P_O(i)\)
\end{minipage} \\
\midrule()
\endhead
6 & 3.1\% & 4.1\% & 4.2\% & 4.3\% & 4.7\% \\
7 & 9.4\% & 11.1\% & 11.2\% & 11.4\% & 11.9\% \\
8 & 16.4\% & 17.7\% & 17.8\% & 17.9\% & 16.5\% \\
9 & 21.9\% & 21.8\% & 21.8\% & 21.8\% & 21.7\% \\
10 & 24.6\% & 23.1\% & 23.0\% & 22.9\% & 21.5\% \\
11 & 24.6\% & 22.1\% & 22.0\% & 21.8\% & 23.7\% \\
\bottomrule()
\end{longtable}

>>>>

\usepackage{longtable}
\begin{document}
\maketitle

\begin{abstract}
My research investigates patterns in round win percentages in
professional Search and Destroy (SnD) matches of the popular
first-person shooter game Call of Duty (CoD). I find evidence in CoD
supprting the hypothesis that round win probability can be modeled as
constant across the series, although not at a naive 50\%. Second, I
examine post-streak round win probability, in search of evidence
positive recency (the ``hot hand'' fallacy) or negative recency (the
``gambler's falacy''). I find that teams perform significantly worse
than expected after streaks of 2, 3, and 4 wins when series end up going
to 9, 10, or 11 (maximum) rounds, suggesting the presence of negative
recency.
\end{abstract}

@@

The empirical offensive round win percentage across all rounds is
47.8\%.\footnote{Offensive round win percentage has been nearly constant
  across the three games during the CDL era: 1. 47.2\% in MW (2020) 2.
  47.9\% in Cold War (2021) 3. 48.1\% in Vanguard (2022)} Table
\ref{tbl:cod-o-win-prop-by-series-state} shows round win percentages by
series ``state'' (i.e.~the number of round wins by each team prior to an
upcoming round). Offensive round win rate is not quite constant,
although never veers more than 10\% from this global average.

@@

Despite the plethora of existing research on games played in a series
and momentum in sports, these topics have yet to be investigated heavily
in esports. Work has been done to examine in-round win probability in
other FPS titles such as Counter-Strike (Xenopoulos, Freeman, and Silva
2022) and Valorant (DeRover 2021), both of which are round-based like
CoD SnD. However, research on round-level trends is sparse, perhaps
because games like Counter-Strike and Valorant both have economic
aspects that can create clear advantages on side in a given round, given
how prior rounds played out.\footnote{Additionally, both Counter-Strike
  and Valorant have overtime rules and blocked offensive/defensive roles
  (i.e.~playing either offense or defense for many consecutive rounds).}

@@

Table \ref{tbl:cod-prob-series-lasting-i-rounds} shows the expected
probabilities for \(s = 11\) under the naive assumption that
\(p_0 = 0.5\) for a given round, along with the observed frequencies
observed in CoD SnD (\(P_O(i)\), \(N_O(i)\))).

\begin{longtable}{rrrr}
  \caption{The probabilities that a best-of-11 series lasts $i$ rounds ($P_E(i)$, where $i = {6, 7, 8, 9, 10, 11}$) under the assumption that each team has a 50% probability ($p_0 = 0.5$) of winning each game. The observed frequencies observed are shown as a count $N_O(i)$ and as a proportion $P_O(i)$ of all series.}\label{tbl:cod-prob-series-lasting-i-rounds} \\
  \toprule
  &  & \multicolumn{2}{c}{Observed} \\ 
  \cmidrule(lr){3-4}
  Series lasts $i$ rounds & $P_E(i)$ & $P_O(i)$ & $N_O(i)$ \\ 
  \midrule
  6 & $3.1\%$ & $4.7\%$ & 40 \\ 
  7 & $9.4\%$ & $11.9\%$ & 101 \\ 
  8 & $16.4\%$ & $16.5\%$ & 141 \\ 
  9 & $21.9\%$ & $21.7\%$ & 185 \\ 
  10 & $24.6\%$ & $21.5\%$ & 183 \\ 
  11 & $24.6\%$ & $23.7\%$ & 202 \\ 
  \bottomrule
\end{longtable}

The probabilities that a best-of-11 series lasts \(i\) rounds (\(P(i)\),
where \(i \in R = [6, 7, 8, 9, 10, 11]\)) under the assumption that each
team has a 50\% probability (\(p_0 = 0.5\)) of winning each game
\{\#tbl-prob-series-lasting-i-rounds\} along with the observed
frequencies of series lasting \(i\) rounds, expressed as a percentage
\(P_O(i)\) and as a count (\(N_O\))

@@

as 16.0 (p-value of 0.0068), I can comfortably reject the constant
probability null hypothesis for \(p_0 = 0.5\), event at a confidence
level of \(\alpha = 0.01\).

@@

Table \ref{tbl:expected-series-lengths-alternative-ps} shows the new
\(P_E(i)\) when re-applying Equation \ref{eq:series-length} for each the
maximum likelihood estimate \(p_2 = 0.575\).\footnote{The method of
  moments and minimum \(\chi^2\) estimates for \(p\) are omitted simply
  because the results would be nearly identical to those for the maximum
  likelihood estimate of \(p\) (since they are all \(\approx 0.575\).} I
observe that \(P_E(i)\) is notably larger for \(p_{2}\) when
\(i \in [6, 7]\), more closely matching \(P_O(i)\). \(p_{2}\) is also
closer to \(P_O(i)\) for \(i \in [9, 10]\), although not for
\(i \in [8, 11]\).

\begin{longtable}[]{@{}rrrr@{}}
\caption{The probabilities that a best-of-11 series lasts $i$ rounds under the naive assumption $p_0 = 0.5$ and the maximum likelihood estimation $p_2 = 0.575$.}\label{tbl:expected-series-lengths-alternative-ps} \\
\toprule()
Series lasts \(i\) rounds & \(p_0\) = 0.5 & \(p_2\) = 0.575 & \(P_O(i)\) \\
\midrule()
\endhead
6 & 3.1\% & 4.2\% & 4.7\% \\
7 & 9.4\% & 11.2\% & 11.9\% \\
8 & 16.4\% & 17.8\% & 16.5\% \\
9 & 21.9\% & 21.8\% & 21.7\% \\
10 & 24.6\% & 23.0\% & 21.5\% \\
11 & 24.6\% & 22.0\% & 23.7\% \\
\bottomrule()
\end{longtable}
